---
title: "Week 1: Loss Distributions Workbook"
format:
  pdf:
    toc: false
---

## Workbook: Exploring the Claim Data

Instructions: Complete the questions by inspecting the provided loss dataset. Show your calculations and answers by hand, then scan or photograph your completed work and submit it as a single PDF file named ID.pdf, where ID is your student ID.


Before fitting probability distributions, explore the claim data and identify their main characteristics.

1. **Inspect the dataset.** How many claim observations are there, and what are the minimum, maximum, mean, and standard deviation of the claim amounts?

2. **Examine the distribution.** Create a histogram of the claim amounts. Does the dataset appear to be symmetric or right-skewed? What features of the histogram suggest that this is an insurance loss dataset?

3. **Examine large claims.** How many claims exceed $5{,}000$ and $10{,}000$, and what percentage of the observations do they represent?

4. **Consider model selection.** Based on your exploration, what characteristics should a probability distribution have to provide a reasonable model for these claim amounts? Which of the exponential, Gamma, lognormal, and Pareto distributions would you consider as potential candidates?


**Solutions**

1. We use R to answer the questions in this workbook. 

```{r}
# Load the claims data. Set the correct path to the data file.

load("~/Documents/Risk_Analysis_Book/data/claims.Rdata")
```

We first transform the loaded data set into a vector of claims, denoted by `claims`.

```{r}
claims <- dat$claims
```

The functions `length(claims)` and `summary(claims)` can be used to inspect the claims data, including the number of observations and descriptive statistics.


```{r}
# number of claims
n <- length(claims)

cat("The number of observations is", n)
```
The number of observations is `r length(claims)` claims.

```{r}
summary(claims)
```
The standard deviation can be obtained by `sd(claims)`.

```{r}
# the standard deviation of claims
s <- sd(claims)

cat("The standard deviation of claims is", s)
```

2. The base R function `hist()` can be used to create a histogram. To set the number of bins in a histogram using the `hist()` function, use the **`breaks`** argument.

You can pass a **single number** to specify the suggested number of bins, or a **numeric vector** to set the exact boundaries of each bin.


```{r}
hist(claims, breaks = 30)
```

From the histogram, the distribution of our synthetic claims dataset appears to be right-skewed. Most claims are relatively small, while a small number of claims are large. These are typical characteristics of loss distributions.

3. There are `r sum(claims > 5000)` (`r 100*sum(claims > 5000)/200`\%) and `r sum(claims > 10000)` (`r 100*sum(claims > 10000)/200`\%) claims that exceed $5{,}000$ and $10{,}000$.

```{r}
# Claims exceeding $5,000
n_5000 <- sum(claims > 5000)
p_5000 <- 100 * n_5000 / length(claims)

# Claims exceeding $10,000
n_10000 <- sum(claims > 10000)
p_10000 <- 100 * n_10000 / length(claims)

cat("Claims exceeding $5,000:", n_5000,
    "(", round(p_5000, 1), "%)\n")

cat("Claims exceeding $10,000:", n_10000,
    "(", round(p_10000, 1), "%)\n")
```

4. Based on the histogram and the characteristics of the claims, distributions such as the lognormal, Pareto, and other fat-tailed distributions are good candidates for modelling the claim data.