---
title: "Workbook 3: Reinsurance and Loss Distributions"
format:
  pdf:
    toc: false
---

## Workbook 3: Reinsurance and Loss Distributions


**Aim:** Use R to explore how reinsurance transforms common loss distributions and how the resulting distributions can be described empirically.

The workbook uses **simulated losses from predefined loss distributions**. It provides **hints throughout**, including both **R code and explanations**, so students can work through the practical step by step.

### Loss Distributions

<!-- 1. **Simulate losses** from Exponential, Gamma, Lognormal, and Pareto II distributions using predefined parameters. -->
1. **Simulate $n=5{,}000$ losses** from the following distributions using the predefined parameters. The parameters are chosen so that all four distributions have approximately the same mean loss, $E(X)=1{,}000$, while exhibiting different shapes and tail behaviour.

| Loss distribution | Distribution and parameters                         |           $E(X)$ |
| ----------------- | --------------------------------------------------- | ---------------: |
| Exponential       | $X\sim\operatorname{Exp}(\lambda=0.001)$            |        $1{,}000$ |
| Gamma             | $X\sim\operatorname{Gamma}(\alpha=2,\lambda=0.002)$ |        $1{,}000$ |
| Lognormal         | $X\sim\operatorname{Lognormal}(\mu=6.408,\sigma=1)$ | $\approx1{,}000$ |
| Pareto            | $X\sim\operatorname{Pa}(\alpha=2,\lambda=1{,}000)$  |        $1{,}000$ |


2. **Explore the simulated losses** using histograms, empirical CDFs, and kernel density estimates.

::: {.callout-note}

## Hint: Empirical CDF

The **empirical cumulative distribution function (ECDF)** describes the proportion of observations that are less than or equal to a given value.

For observations $x_1,\ldots,x_n$,

$$
\widehat{F}_n(x)
=
\frac{1}{n}\sum_{i=1}^n I(x_i \leq x).
$$

**What is it for?**
The ECDF provides a direct, non-parametric view of the distribution and can be used to estimate probabilities and compare distributions.

**In R:** use `ecdf()` to create the ECDF function and `plot()` to display it.

```r
F_emp <- ecdf(losses)
plot(F_emp)
```

Students can run the code by placing the cursor in the code cell and selecting **Run** in RStudio, or by clicking the green ▶ button beside the code chunk.
:::

::: {.callout-note}

## Hint: Kernel Density Estimate

A **kernel density estimate (KDE)** provides a smooth estimate of the probability density from observed data.

**What is it for?**
KDE helps visualize the shape of a distribution without assuming a particular parametric distribution.

**In R:** use `density()` to calculate the estimate and `plot()` to display it.

```r
dens <- density(losses)
plot(dens)
```

The choice of bandwidth controls the amount of smoothing: a smaller bandwidth produces a more detailed curve, while a larger bandwidth produces a smoother curve.
:::

### XOL Reinsurance

3. **Apply XOL reinsurance** with no deductible and limit $M=2{,}000$:
    
    $$
    Y=\min(X,M).
    $$
    
    For each of the four simulated loss distributions:

    * calculate the XOL payments $Y$;
    * create a histogram of the payments;
    * plot the empirical CDF of $Y$; and
    * identify and calculate the probability mass at $M$.    

### Reinsurer's XOL Payment
  
4. **Calculate the reinsurer's XOL payment** with attachment point $M=2{,}000$:

    $$
    Z=(X-M)^+=\max(X-M,0).
    $$

    For each loss distribution:
    
    * calculate the reinsurer's payments $Z$;
    * plot the histogram;
    * plot the ECDF; and
    * calculate the probability mass at zero.
  
5. **Investigate the mixed distribution.** Explain why the reinsurer's payment has a point mass at zero and how this is reflected in the ECDF.

### Compare and Interpret

6. **Compare the insurer's retained payments and the reinsurer's payments** across the four loss distributions. Consider:

    * the mean payment;
    * the variability of payments; and
    * the probability of very large payments.

7. **Interpret the results.** How does XOL reinsurance change the direct insurer's **mean payment, variability, and exposure to very large claims**?
  
    * the **mean amount paid** by the direct insurer;
    * the **variability of the amount paid**; and
    * the **probability that the direct insurer pays a very large claim.


::: {.callout-note}

## Hint: Plotting Multiple Curves with `ggplot2`

When comparing several distributions, `ggplot2` can display multiple curves in the same plot. Map the distribution name to `colour` to automatically obtain different colours and a legend.

For example, suppose the data are stored in a data frame with `loss` and `distribution`:
  
```r
library(ggplot2)

ggplot(data, aes(x = loss, colour = distribution)) +
  stat_ecdf() +
  labs(
    x = "Loss",
    y = "Empirical CDF",
    colour = "Distribution"
  )
```

The key part is:
  
```r
aes(colour = distribution)
```

This tells `ggplot2` to use a different colour for each distribution and to create a corresponding legend.

The same approach can be used for kernel density estimates:
  
```r
ggplot(data, aes(x = loss, colour = distribution)) +
  geom_density() +
  labs(
    x = "Loss",
    y = "Density",
    colour = "Distribution"
  )
```

**Tip:** Keep the same colour mapping when comparing different plots. This makes it easier to identify each distribution across the workbook.
:::
 

::: {.callout-note}

## Hint: Useful `stat_*` Functions

For this practical, the main `stat_*` function is:
  
  * **`stat_ecdf()`** — calculates and displays the empirical cumulative distribution function from the observed data.

```r
ggplot(data, aes(x = loss, colour = distribution)) +
  stat_ecdf()
```

This produces an empirical CDF curve for each distribution.

For kernel density estimation, use **`density()`** to calculate the KDE and then display the result with `geom_line()` or `geom_density()` when appropriate.

For example:
  
```r
dens <- density(losses)

ggplot(data.frame(x = dens$x, y = dens$y),
       aes(x = x, y = y)) +
  geom_line()
```

For this workbook, `stat_ecdf()` is particularly useful because the ECDF can represent **both continuous and mixed distributions**, including the point mass created by XOL.
:::


<!-- 4. **Explore the XOL payment distribution** using the ECDF and KDE. -->

<!-- 5. **Investigate the mixed distribution**, paying particular attention to the probability mass at $M$. -->

<!-- ### Proportional Reinsurance -->

<!-- 4. **Apply proportional reinsurance** with retention $c$: -->

<!-- $$ -->
<!-- Y = cX. -->
<!-- $$ -->

<!-- 5. **Compare the original and proportional-reinsurance distributions** using the ECDF and KDE. -->

<!-- ### Compare and Interpret -->

<!-- 6. **Compare XOL and proportional reinsurance** across the four loss distributions. -->

<!-- 7. **Interpret the results:** How does each reinsurance arrangement change the distribution? Why does XOL produce a mixed distribution, whereas proportional reinsurance simply rescales the losses? -->


**Solutons**

1.**Simulate $n=5{,}000$ losses** from the following distributions using the predefined parameters.

The R code for simulating the Exponential distribution.

```{r}
set.seed(123)

n <- 5000
lambda <- 0.001

loss_exp <- rexp(n, rate = lambda)
```

The simulated losses are stored in `loss_exp`, with

$$
X \sim \operatorname{Exp}(\lambda=0.001),
\qquad E(X)=\frac{1}{\lambda}=1{,}000.
$$


2. **Explore the simulated Exponential losses** using a histogram, empirical CDF, and kernel density estimate.

**Histogram**

```{r}
library(ggplot2)

ggplot(data.frame(loss = loss_exp), aes(x = loss)) +
  geom_histogram(bins = 50) +
  labs(x = "Loss", y = "Count")
```

**Empirical CDF**

```{r}
ggplot(data.frame(loss = loss_exp), aes(x = loss)) +
  stat_ecdf() +
  labs(x = "Loss", y = "Empirical CDF")
```

**Kernel density estimate**

```{r}
ggplot(data.frame(loss = loss_exp), aes(x = loss)) +
  geom_density() +
  labs(x = "Loss", y = "Density")
```

These three plots provide complementary views of the simulated loss distribution: the **histogram** shows the observed frequencies, the **ECDF** shows cumulative probabilities, and the **KDE** provides a smooth estimate of the density.


3. **Apply XOL reinsurance** with no deductible and limit $M=2{,}000$:

```{r}
# XOL parameters
M <- 2000

# XOL payments
xol_exp <- pmin(loss_exp, M)

# Histogram
ggplot(data.frame(payment = xol_exp), aes(x = payment)) +
  geom_histogram(bins = 50) +
  labs(x = "XOL payment", y = "Count")

# Empirical CDF
ggplot(data.frame(payment = xol_exp), aes(x = payment)) +
  stat_ecdf() +
  labs(x = "XOL payment", y = "Empirical CDF")

# Probability mass at M
mean(xol_exp == M)
```

The last calculation estimates

$$
P(Y=M)=P(X\geq M),
$$

which is the **point mass at the XOL limit**.


**Expected results**

* **Histogram:** Most payments are below $2{,}000$, with a clear concentration of observations at the XOL limit $M=2{,}000$.
* **ECDF:** The curve shows a **jump at $2{,}000$**. The size of this jump is approximately $0.1352$, representing the probability mass at the limit.
* **KDE:** A standard KDE of all payments is **not appropriate** for displaying the mixed distribution because it does not represent the point mass at $M$ correctly.
* **Probability mass:** Approximately **13.52%** of the XOL payments are exactly $2{,}000$:

$$
P(Y=2{,}000)\approx0.1352.
$$

4. **Calculate the reinsurer's XOL payment** with attachment point $M=2{,}000$:

For the reinsurer's XOL payment with attachment point $M=2{,}000$,

$$
Z=(X-M)^+=\max(X-M,0).
$$

```{r}
# XOL attachment point
M <- 2000

# Reinsurer's XOL payment
xol_reinsurer_exp <- pmax(loss_exp - M, 0)

# Histogram
ggplot(data.frame(payment = xol_reinsurer_exp), aes(x = payment)) +
  geom_histogram(bins = 50) +
  labs(x = "Reinsurer's XOL payment", y = "Count")

# Empirical CDF
ggplot(data.frame(payment = xol_reinsurer_exp), aes(x = payment)) +
  stat_ecdf() +
  labs(x = "Reinsurer's XOL payment", y = "Empirical CDF")

# Probability mass at zero
mean(xol_reinsurer_exp == 0)
```

The last line estimates

$$
P(Z=0)=P(X\leq M),
$$

which is the **point mass at zero** in the reinsurer's payment distribution.


5. **Investigate the mixed distribution.** Explain why the reinsurer's payment has a point mass at zero and how this is reflected in the ECDF.

The reinsurer pays nothing when $X\leq M$, so there is a point mass at zero:

$$
P(Z=0)=P(X\leq M).
$$

In the ECDF, this appears as a **vertical jump at $Z=0$**.


6. **Compare the insurer's retained payments and the reinsurer's payments** across the four loss distributions. Consider the mean payment, variability, and probability of very large payments.

For the Exponential losses:

```{r} 
data.frame(
  Payment = c("Insurer", "Reinsurer"),
  Mean = c(mean(xol_exp), mean(xol_reinsurer_exp)),
  SD = c(sd(xol_exp), sd(xol_reinsurer_exp)),
  P_Large = c(
    mean(xol_exp > 2000),
    mean(xol_reinsurer_exp > 2000)
  )
)
```

You can record the results for all four distributions in the following table:

| Distribution | Insurer Mean | Insurer SD | Insurer $P(\text{large})$ | Reinsurer Mean | Reinsurer SD | Reinsurer $P(\text{large})$ |
| ------------ | -----------: | ---------: | ------------------------: | -------------: | -----------: | --------------------------: |
| Exponential  |              |            |                           |                |              |                             |
| Gamma        |              |            |                           |                |              |                             |
| Lognormal    |              |            |                           |                |              |                             |
| Pareto       |              |            |                           |                |              |                             |

Here, define a **large payment** consistently across all distributions, for example, as a payment exceeding $2{,}000$.

7. **Interpret the results.** How does XOL reinsurance change the direct insurer's **mean payment, variability, and exposure to very large claims**?

XOL reinsurance **reduces the direct insurer's exposure to large losses**. Compared with the original losses:

* **Mean payment:** decreases because losses above the attachment point are transferred to the reinsurer.
* **Variability:** decreases because the insurer's payment is capped at the XOL limit.
* **Very large claims:** the probability of the direct insurer paying a very large amount is substantially reduced because payments are limited by the reinsurance arrangement.
