---
title: "Chapter 3: Risk Models"

format:
  beamer:
    theme: Boadilla
    colortheme: default
    fonttheme: professionalfonts
    aspectratio: 169
    pdf-engine: xelatex
    keep-tex: true
    toc: false
    number-sections: false
    
# format:
#   revealjs:
#     theme: simple
#     css: styles.css
#     slide-number: true
#     scrollable: true
#     chalkboard:
#       buttons: false
#     preview-links: auto

execute:
  echo: false

---


# Compound Negative Binomial Distributions {#compound-negative-binomial}

## From Compound Poisson to Compound Negative Binomial

In the collective risk model,

$$
S=\sum_{i=1}^{N}X_i.
$$

The distribution of $S$ depends on both:

- the claim number distribution $N$;
- the claim severity distribution $X$.

So far, we have considered the compound Poisson model.

## Why Do We Need Another Frequency Distribution?

For a Poisson random variable,

$$
E[N]=\operatorname{Var}(N)=\lambda.
$$

However, claim-count data may exhibit **overdispersion**:

$$
\operatorname{Var}(N)>E[N].
$$

In such cases, the Poisson distribution may be too restrictive.

The negative binomial distribution provides an alternative.

## The Negative Binomial Distribution — Type 2

Under the IFoA Type 2 convention,

$$
N\sim NB(k,p),
$$

where $N$ is the number of **failures before the $k$th success**.

The parameters satisfy

$$
k>0,\qquad 0<p<1,\qquad q=1-p.
$$

The support is

$$
N=0,1,2,\ldots
$$

## What Does the Negative Binomial Count?

Imagine a sequence of independent Bernoulli trials.

We continue until the $k$th success occurs.

The random variable $N$ counts the number of failures observed before that success.

For example, when $k=3$,

$$
F,F,S,F,F,S,F,S
$$

gives $N=5$.

## Probability Function

For $n=0,1,2,\ldots$,

$$
\Pr(N=n)
=
\frac{\Gamma(k+n)}
{\Gamma(n+1)\Gamma(k)}
p^kq^n.
$$

When $k$ is an integer,

$$
\Pr(N=n)
=
\binom{k+n-1}{n}p^kq^n.
$$

This is the **Negative Binomial Type 2** parameterisation.

## Properties of the Negative Binomial

The mean is

$$
E[N]=\frac{kq}{p}.
$$

The variance is

$$
\operatorname{Var}(N)=\frac{kq}{p^2}.
$$

Since $0<p<1$,

$$
\operatorname{Var}(N)>E[N].
$$

## Probability Generating Function

The probability generating function is

$$
G_N(s)
=
E[s^N]
=
\left(\frac{p}{1-qs}\right)^k.
$$

The moment generating function is obtained by setting

$$
s=e^t.
$$

Hence,

$$
M_N(t)
=
\left(\frac{p}{1-qe^t}\right)^k.
$$

## Poisson versus Negative Binomial

| | Poisson | Negative Binomial |
|---|---|---|
| Mean | $\lambda$ | $kq/p$ |
| Variance | $\lambda$ | $kq/p^2$ |
| Dispersion | $\operatorname{Var}(N)=E[N]$ | $\operatorname{Var}(N)>E[N]$ |

The negative binomial allows **greater variability in claim counts**.

## Why Use the Negative Binomial for Claim Counts?

Insurance portfolios contain risks with different claim frequencies.

Consequently, observed claim counts may be more variable than allowed by a Poisson model.

The negative binomial provides a flexible frequency model with

$$
\operatorname{Var}(N)>E[N].
$$

This makes it useful for modelling **overdispersed claim counts**.

## From Frequency to Aggregate Loss

The collective risk model is

$$
S=\sum_{i=1}^{N}X_i.
$$

Now let the claim number distribution be

$$
N\sim NB(k,p).
$$

The aggregate loss $S$ is therefore a **compound negative binomial random variable**.

## The Compound Negative Binomial Distribution

Let

$$
N\sim NB(k,p)
$$

and let

$$
X_1,X_2,\ldots\sim F_X
$$

be independent and identically distributed claim amounts, independent of $N$.

Then

$$
S=\sum_{i=1}^{N}X_i
$$

has a **compound negative binomial distribution**.

## Notation

We denote the compound negative binomial distribution by

$$
S\sim\mathcal{CNB}(k,p,F_X).
$$

Here,

- $k,p$ determine the claim frequency distribution;
- $F_X$ determines the claim severity distribution;
- $S$ represents aggregate claims.

The model separates **frequency** from **severity**.

## Mean of the Compound Negative Binomial

For a compound distribution,

$$
E[S]=E[N]E[X].
$$

Let $m_1=E[X].$

Since

$$
E[N]=\frac{kq}{p},
$$

we obtain

$$
\boxed{E[S]=\frac{kq}{p}m_1}.
$$

## Variance of the Compound Negative Binomial

For a compound distribution,

$$
\operatorname{Var}(S)
=
E[N]\operatorname{Var}(X)
+
\operatorname{Var}(N)(E[X])^2.
$$

Let

$$
m_1=E[X],
\qquad
m_2=E[X^2].
$$

Then

$$
\boxed{
\operatorname{Var}(S)
=
\frac{kq}{p^2}
\left(pm_2+qm_1^2\right)
}.
$$

## Interpreting the Variance

The aggregate loss variance reflects two sources of uncertainty:

$$
\operatorname{Var}(S)
=
\underbrace{E[N]\operatorname{Var}(X)}
_{\text{severity variation}}
+
\underbrace{\operatorname{Var}(N)(E[X])^2}
_{\text{frequency variation}}.
$$

Thus, both claim frequency and claim severity contribute to aggregate risk.

## MGF of the Compound Distribution

For a compound distribution,

$$
M_S(t)=G_N(M_X(t)).
$$

For the negative binomial,

$$
G_N(s)
=
\left(\frac{p}{1-qs}\right)^k.
$$

Therefore,

$$
\boxed{
M_S(t)
=
\left(
\frac{p}{1-qM_X(t)}
\right)^k
}.
$$

## Worked Example

Suppose $N\sim NB(k,p),$ with $k=4,\qquad p=0.5.$

Claim amounts have $m_1=E[X]=1000,\qquad m_2=E[X^2]=1{,}500{,}000.$

Calculate:

* the expected number of claims;
* the variance of the number of claims;
* the expected aggregate loss;
* the variance of aggregate loss.

## Worked Example: Claim Number

Given $k=4$, $p=0.5$, and $q=0.5$,

$$
E[N]=\frac{kq}{p}
=\frac{4(0.5)}{0.5}=4.
$$

$$
\operatorname{Var}(N)
=\frac{kq}{p^2}
=\frac{4(0.5)}{0.5^2}=8.
$$

## Worked Example: Aggregate Loss

Given $m_1=1{,}000$ and $m_2=1{,}500{,}000$,

$$
E[S]=\frac{kq}{p}m_1
=4(1{,}000)=4{,}000.
$$

$$
\operatorname{Var}(S)
=\frac{kq}{p^2}(pm_2+qm_1^2)
=8[0.5(1{,}500{,}000)+0.5(1{,}000)^2]
=10{,}000{,}000.
$$




## R Simulation

We can simulate the collective risk model in two stages.

First, simulate the number of claims:

$$
N\sim NB(k,p).
$$

Then simulate the claim amounts

$$
X_1,\ldots,X_N\sim F_X.
$$

Finally, calculate

$$
S=\sum_{i=1}^{N}X_i.
$$





## R Simulation: Setting Up the Model

We simulate $n_{\text{sim}}=5{,}000$ insurance years.

```r
set.seed(123)

k <- 4
p <- 0.5
q <- 1 - p

shape <- 2
rate <- 0.002
n_sim <- 5000
```

Here, `k` and `p` define the negative binomial frequency, while `shape` and `rate` define the Gamma severity distribution.

## R Simulation: Claim Numbers

First, simulate the number of claims in each year.

```r
N <- rnbinom(
  n_sim,
  size = k,
  prob = p
)
```

Each element of `N` represents the number of claims in one simulated year.

Thus,

$$
N\sim NB(k,p).
$$

For these parameters,

$$
E[N]=4,\qquad \operatorname{Var}(N)=8.
$$

## R Simulation: Claim Amounts

For each simulated year, generate the individual claim amounts.

```r
S <- sapply(N, function(n) {
  if (n == 0) return(0)
  sum(rgamma(
    n, shape = shape, rate = rate
  ))
})
```

For a year with $N=n$ claims, this code generates $n$ Gamma claim amounts and adds them together.

Thus,

$$
S=\sum_{i=1}^{N}X_i.
$$

## Claim Number: Simulation Result

```r
hist(
  N,
  breaks = seq(-0.5, max(N) + 0.5, 1),
  main = "Claim Number: Negative Binomial",
  xlab = "Number of claims, N"
)
```

The histogram shows the simulated distribution of the annual claim count.

Each bar represents the number of simulated years having a particular value of $N$.

The distribution includes the possibility of **zero claims**.

## Aggregate Loss: Simulation Result

```r
hist(
  S,
  breaks = 50,
  main = "Aggregate Loss: Compound Negative Binomial",
  xlab = "Aggregate loss, S"
)
```

The histogram shows the simulated distribution of annual aggregate losses.

Each value of `S` represents one simulated insurance year.

The right-skewed shape arises from variation in both **claim frequency** and **claim severity**.

## Simulation: From Frequency to Aggregate Loss

The simulation follows the compound model in three stages:

$$
N\sim NB(k,p)
$$

$$
X_1,\ldots,X_N\sim F_X
$$

$$
S=\sum_{i=1}^{N}X_i.
$$

Thus, each element of `S` represents **one simulated aggregate loss**.

## Simulation versus Theory

We can calculate the theoretical severity moments.

```r
m1 <- shape / rate
m2 <- shape * (shape + 1) / rate^2

theory_mean <- k * q / p * m1
theory_var <- k * q / p^2 *
  (p * m2 + q * m1^2)
```

Here, `m1` and `m2` are the first and second moments of the Gamma severity.

These are substituted into the compound negative binomial formulas.

## Simulation versus Theory

Compare the simulated and theoretical results.

```r
c(
  Simulated_Mean = mean(S),
  Theoretical_Mean = theory_mean,
  Simulated_Var = var(S),
  Theoretical_Var = theory_var
)
```

The simulated values should be close to the theoretical values.

With many simulated years, the differences become smaller.

This provides a numerical check of the compound negative binomial formulas.

## Theoretical Aggregate Loss

For the Gamma severity distribution,

$$
E[X]=\frac{\text{shape}}{\text{rate}}=1000
$$

and

$$
E[X^2]
=
\frac{\text{shape}(\text{shape}+1)}
{\text{rate}^2}
=1{,}500{,}000.
$$

Therefore,

$$
E[S]
=
\frac{kq}{p}m_1
=
4(1000)
=
4000.
$$

The simulation should produce an aggregate mean close to $4000$.


## Graph: Aggregate Loss {.smaller}

```{r}
#| fig-width: 6
#| fig-height: 3.5

set.seed(123)

n_sim <- 5000
k <- 4
p <- 0.5

shape <- 2
rate <- 0.002

N <- rnbinom(
  n_sim, size = k, prob = p
)

S <- sapply(N, function(n) {
  if (n == 0) return(0)
  sum(rgamma(
    n, shape = shape, rate = rate
  ))
})

hist(
  S, breaks = 50,
  main = "Aggregate Loss: Compound Negative Binomial",
  xlab = "Aggregate loss, S",
  ylab = "Frequency"
)
```



## Compare Simulation with Theory

And similarly, this can be a **self-contained chunk** that calculates the theoretical moments and compares them with the simulation:

```{r}
m1 <- shape / rate
m2 <- shape * (shape + 1) / rate^2

theory_mean <- k * (1 - p) / p * m1
theory_var <- k * (1 - p) / p^2 *
  (p * m2 + (1 - p) * m1^2)

data.frame(
  Measure = c("Mean", "Variance"),
  Simulation = c(mean(S), var(S)),
  Theory = c(theory_mean, theory_var)
)
```



## Interpreting the Simulation

The simulated and theoretical results are reasonably close:

| Measure  |  Simulation |     Theory |
| -------- | ----------: | ---------: |
| Mean     |     3,926.8 |      4,000 |
| Variance | 9,437,849.2 | 10,000,000 |

* The simulated mean is close to the theoretical mean of $4,000$.
* The simulated variance is also reasonably close to the theoretical value.
* The differences arise because only $5,000$ insurance years were simulated.
* Increasing the number of simulations would generally make the simulated results closer to the theoretical values.
* This provides a numerical check of the compound negative binomial formulas.

The key idea is that **simulation reproduces the theoretical behaviour of the model approximately**.



## When Might We Use This Model?

The compound negative binomial model is useful when:

- claim counts are overdispersed;
- $\operatorname{Var}(N)>E[N]$;
- the Poisson model is too restrictive;
- a flexible frequency model is required;
- aggregate claim amounts are the quantity of interest.

## Limitations

The negative binomial distribution is not automatically appropriate whenever

$$
\operatorname{Var}(N)>E[N].
$$

The choice of frequency distribution should be supported by:

- the characteristics of the data;
- goodness-of-fit analysis;
- the modelling assumptions;
- the intended application.

## Summary

The key progression is

$$
\text{Poisson}
\rightarrow
\text{Overdispersion}
\rightarrow
\text{Negative Binomial}
\rightarrow
\text{Compound Negative Binomial}.
$$

For $N\sim NB(k,p)$,

$$
E[N]=\frac{kq}{p},
\qquad
\operatorname{Var}(N)=\frac{kq}{p^2}.
$$

## Summary: Aggregate Loss

For $S=\sum_{i=1}^{N}X_i,$

the compound negative binomial model gives

$$
E[S]=\frac{kq}{p}m_1,
$$

and

$$
\operatorname{Var}(S)
=
\frac{kq}{p^2}
\left(pm_2+qm_1^2\right).
$$

Its MGF is

$$
M_S(t)
=
\left(
\frac{p}{1-qM_X(t)}
\right)^k.
$$












# Compound Binomial Distributions {#compound-binomial}

## From Individual Policies to Aggregate Loss

Consider a portfolio of $n$ independent and identical policies.

Each policy can give rise to **at most one claim** during the period.

Let $p$ be the probability that a policy produces a claim.

The number of claims in the portfolio is therefore

$$
N \sim \operatorname{Binomial}(n,p).
$$

## Why the Binomial Distribution?

The binomial model is natural when:

* there are a fixed number $n$ of potential claim events;
* each policy produces either **zero or one claim**;
* each policy has the same claim probability $p$;
* claim events are independent.

Thus, the number of claims has an upper limit:

$$
0 \leq N \leq n.
$$

## The Compound Binomial Model

Let $X_i$ denote the amount of the $i$th claim.

Assume that $X_1,X_2,\ldots$ are i.i.d. and independent of $N$.

Then the aggregate claim amount is

$$
S=\sum_{i=1}^{N}X_i.
$$

We call $S$ a **compound binomial random variable**.

The distribution is denoted by

$$
S\sim\mathcal{CB}(n,p,F_X).
$$

## Aggregate Loss Model

The collective risk model is

$$
S=\sum_{i=1}^{N}X_i,
$$

where

$$
N\sim\operatorname{Binomial}(n,p).
$$

Here, $N$ determines **how many claims occur**, while $X_i$ determines **how much each claim costs**.

Thus, the aggregate loss combines a **frequency model** and a **severity model**.

## Expected Aggregate Loss

For a random sum,

$$
S=\sum_{i=1}^{N}X_i,
$$

the expected aggregate loss is

$$
\boxed{E[S]=E[N]E[X]}.
$$

This follows from the standard result for a random sum when $N$ is independent of the claim amounts.

## Binomial Frequency Model

For the binomial frequency model,

$$
N\sim\operatorname{Binomial}(n,p),
$$

so

$$
E[N]=np.
$$

Let

$$
m_1=E[X].
$$

Substituting into the random-sum formula,

$$
\begin{aligned}
E[S]
&=E[N]E[X]\\
&=(np)m_1.
\end{aligned}
$$

Therefore,

$$
\boxed{E[S]=npm_1}.
$$


## Interpretation of Expected Aggregate Loss

The formula

$$
E[S]=npm_1
$$

can be written as

$$
\boxed{
E[S]
=
\underbrace{np}_{\text{expected number of claims}}
\underbrace{m_1}_{\text{expected claim amount}}
}.
$$

Thus, expected aggregate loss is determined by:

* the expected number of claims, $np$;
* the expected amount per claim, $m_1$.

The binomial frequency model therefore gives a simple interpretation of the expected aggregate loss.

## Variance of Aggregate Loss

For a random sum,

$$
\operatorname{Var}(S)
=
E[N]\operatorname{Var}(X)
+
\operatorname{Var}(N)\{E[X]\}^2.
$$

For

$$
N\sim\operatorname{Binomial}(n,p),
$$

we have

$$
E[N]=np,
\qquad
\operatorname{Var}(N)=np(1-p).
$$

Let

$$
m_1=E[X],
\qquad
m_2=E[X^2].
$$

## Substituting the Binomial Moments

Since

$$
\operatorname{Var}(X)=m_2-m_1^2,
$$

substitute into the random-sum formula:

$$
\begin{aligned}
\operatorname{Var}(S)
&=np(m_2-m_1^2)
+np(1-p)m_1^2.
\end{aligned}
$$

Expanding,

$$
\begin{aligned}
\operatorname{Var}(S)
&=npm_2-npm_1^2
+np(1-p)m_1^2.
\end{aligned}
$$

## Simplifying the Variance

Collect the terms involving $m_1^2$:

$$
\begin{aligned}
\operatorname{Var}(S)
&=npm_2
-npm_1^2
+np(1-p)m_1^2\\
&=npm_2-np^2m_1^2.
\end{aligned}
$$

Therefore,

$$
\boxed{
\operatorname{Var}(S)
=
npm_2-np^2m_1^2
}.
$$

The result combines **claim-size uncertainty** and **frequency uncertainty** under the binomial model.


## Worked Example: Compound Binomial with Gamma Severity

Suppose

$$
N\sim\operatorname{Binomial}(100,0.01),
$$

and, independently,

$$
X_i\overset{\text{i.i.d.}}{\sim}\operatorname{Gamma}(20,0.1),
$$

where the Gamma distribution uses the **shape-rate** parameterisation.

Let

$$
S=\sum_{i=1}^{N}X_i.
$$

## Worked Example: Tasks

For the compound binomial model above, find:

1. the MGF of $S$;
2. $E[S]$ using the MGF;
3. $\operatorname{Var}(S)$ using the MGF;
4. $P(S=0)$.


## 1. MGF of the Aggregate Loss

For

$$
X\sim\operatorname{Gamma}(20,0.1),
$$

the MGF is

$$
M_X(t)
=
\left(\frac{0.1}{0.1-t}\right)^{20}.
$$

For the compound binomial model,

$$
M_S(t)
=
\left[q+pM_X(t)\right]^n.
$$

Here, $n=100,\qquad p=0.01,\qquad q=0.99.$

## MGF of $S$

Substituting the values,

$$
\boxed{
M_S(t)=
\left[
0.99+
0.01\left(\frac{0.1}{0.1-t}\right)^{20}
\right]^{100}
}
$$

for $t<0.1$.

This follows directly from the compound binomial MGF.

## 2. Expected Aggregate Loss

Use $E[S]=M_S'(0).$

Let

$$
A(t)=
0.99+
0.01\left(\frac{0.1}{0.1-t}\right)^{20}.
$$

Then

$$
M_S(t)=A(t)^{100}.
$$

Hence, $M_S'(t)=100A(t)^{99}A'(t).$

At $t=0$, \qquad $A(0)=1, \qquad A'(0)=0.01M_X'(0).$

## Expected Aggregate Loss

For the Gamma distribution,

$$
M_X'(0)=E[X]
=\frac{20}{0.1}=200.
$$

Therefore,

$$
\begin{aligned}
E[S]
&=M_S'(0)\\
&=100(0.01)(200).
\end{aligned}
$$

Thus,

$$
\boxed{E[S]=200}.
$$


## 3. Variance Using the MGF

Use

$$
\operatorname{Var}(S)
=
M_S''(0)-[M_S'(0)]^2.
$$

From the previous result,

$$
M_S'(0)=200.
$$

We therefore need

$$
M_S''(0).
$$

## Second Derivative

For

$$
M_S(t)=A(t)^{100},
$$

the second derivative is

$$
M_S''(t)
=
100(99)A(t)^{98}[A'(t)]^2
+
100A(t)^{99}A''(t).
$$

At $t=0$,

$$
M_S''(0)
=
100(99)[A'(0)]^2
+
100A''(0).
$$

## Evaluate the Components

For the Gamma severity,

$$
E[X^2]
=
\operatorname{Var}(X)+[E(X)]^2.
$$

Since

$$
\operatorname{Var}(X)=\frac{20}{0.1^2}=2{,}000,
$$

we obtain

$$
E[X^2]=2{,}000+200^2=42{,}000.
$$

Therefore, $A'(0)=0.01(200), \qquad A''(0)=0.01(42{,}000).$

## Variance of $S$

Thus,

$$
M_S''(0)
=
100(99)(0.01\times200)^2
+
100(0.01\times42{,}000).
$$

Hence,

$$
M_S''(0)=81{,}600.
$$

Finally,

$$
\begin{aligned}
\operatorname{Var}(S)
&=81{,}600-200^2\\
&=\boxed{41{,}600}.
\end{aligned}
$$

## 4. Probability of Zero Aggregate Loss

Because claim amounts are positive,

$$
S=0
\quad\Longleftrightarrow\quad
N=0.
$$

Therefore,

$$
P(S=0)=P(N=0).
$$

For

$$
N\sim\operatorname{Binomial}(100,0.01),
$$

we have

$$
P(N=0)
=
(1-0.01)^{100}.
$$

## Evaluating $P(S=0)$

Thus,

$$
P(S=0)
=
(0.99)^{100}.
$$

Numerically,

$$
\boxed{P(S=0)\approx0.3660}.
$$

So there is approximately a **36.6% probability of no aggregate claims**.

## An Alternative MGF View

Recall that $M_S(t)=E[e^{tS}].$

As $t\to-\infty$,

$$
e^{tS}\to
\begin{cases}
1, & S=0,\\
0, & S>0.
\end{cases}
$$

Therefore,

$$
P(S=0)
=
\lim_{t\to-\infty}M_S(t).
$$

For this model, $\qquad \lim_{t\to-\infty}M_S(t)=(0.99)^{100} \approx0.3660.$

This provides another useful interpretation of the MGF.

## Key Lessons

The compound binomial MGF is obtained by substituting the severity MGF into

$$
M_S(t)=\left[q+pM_X(t)\right]^n.
$$

For this example, $\boxed{M_S(t)=\left[0.99+0.01\left(\frac{0.1}{0.1-t}\right)^{20}\right]^{100}}.$

The MGF can be used to obtain moments:

$$
E[S]=200,
\qquad
\operatorname{Var}(S)=41{,}600.
$$

The model has a positive probability of zero aggregate loss: $P(S=0)=(0.99)^{100}\approx0.3660.$

**Main idea:** the compound-binomial MGF provides a single framework for obtaining both **aggregate-loss moments** and other distributional information.








## R Simulation

We can simulate the collective risk model in two stages.

1. **Simulate the number of claims** $N$.
2. **Simulate the claim amounts** $X_1,\ldots,X_N$ and calculate the aggregate loss

   $$
   S=\sum_{i=1}^{N}X_i.
   $$

For this example,

$$
N\sim\operatorname{Binomial}(100,0.01)
$$

and

$$
X_i\overset{\text{i.i.d.}}{\sim}\operatorname{Gamma}(20,0.1).
$$

## Stage 1: Simulate the Number of Claims

For each simulated year, generate

$$
N\sim\operatorname{Binomial}(100,0.01).
$$

```{r}
set.seed(123)

B <- 100000

N <- rbinom(
  B, size = 100, prob = 0.01
)
```

Each element of `N` represents the **number of claims in one simulated year**.

For example,

```{r}
head(N)
```

The simulation produces $B=100{,}000$ independent insurance years.

## Stage 2: Simulate Aggregate Loss

For each simulated year:

* If $N=0$, then $S=0$.
* If $N=n>0$, simulate $n$ claim amounts.
* Add the claim amounts to obtain $S$.

```{r}
S <- sapply(N, function(n) {
  if (n == 0) {
    0
  } else {
    sum(rgamma(
      n, shape = 20, rate = 0.1
    ))
  }
})
```

Thus, each element of `S` represents one simulated value of

$$
S=\sum_{i=1}^{N}X_i.
$$


## R Simulation

We can simulate the aggregate loss directly in R.

```r
# Simulate aggregate losses
set.seed(123)
B <- 100000
N <- rbinom(B, size = 100, prob = 0.01)
S <- sapply(N, function(n) {
  sum(rgamma(n, shape = 20, rate = 0.1))
})

# Plot the simulated aggregate losses
hist(S, breaks = 50,
     main = "Simulated Aggregate Loss",
     xlab = "Aggregate loss, S")
```

## R Simulation

We can simulate the aggregate loss directly in R.

```{r}
#| echo: false
#| fig-width: 5.5
#| fig-height: 3.5
#| fig-align: "center"

# Simulate aggregate losses
set.seed(123)
B <- 100000
N <- rbinom(B, size = 100, prob = 0.01)
S <- sapply(N, function(n) {
  sum(rgamma(n, shape = 20, rate = 0.1))
})

# Plot the simulated aggregate losses
hist(S, breaks = 50,
     main = "Simulated Aggregate Loss",
     xlab = "Aggregate loss, S",
     col = "lightblue",
     border = "white")
```


## Simulation Results

Use the simulated aggregate losses to estimate the moments.

```{r}
mean(S)
var(S)
```

For $100{,}000$ simulations, $\widehat{E[S]}=199.89$ and $\widehat{\operatorname{Var}(S)}=41{,}515.39$.


Compare with the theoretical results:

$$
E[S]=200,
\qquad
\operatorname{Var}(S)=41{,}600.
$$

**The simulation closely reproduces the theoretical results.**

## From Model to Simulation

The simulation follows exactly the structure of the collective risk model:

$$
\boxed{
N
\longrightarrow
X_1,\ldots,X_N
\longrightarrow
S
}
$$
**Frequency:** $N\sim\operatorname{Binomial}(100,0.01)$

**Severity:** $X_i\sim\operatorname{Gamma}(20,0.1)$

**Aggregate loss:** $S=\sum_{i=1}^{N}X_i$


> **Key idea:** simulation lets us generate many possible insurance years and use the simulated aggregate losses to approximate quantities such as $E[S]$ and $\operatorname{Var}(S)$.










# Properties of the Compound Binomial Model

## Comparing the Three Frequency Models

The three models differ mainly in their assumptions about the number of claims $N$.

| Model                      | Frequency distribution            | Support of $N$ |
| -------------------------- | --------------------------------- | -------------- |
| Compound Binomial          | $\operatorname{Binomial}(n,p)$    | $0,\ldots,n$   |
| Compound Poisson           | $\operatorname{Poisson}(\lambda)$ | $0,1,2,\ldots$ |
| Compound Negative Binomial | $\operatorname{NB}(k,p)$          | $0,1,2,\ldots$ |

The claim amounts $X_i$ can have the same severity distribution in all three models.

## Comparing Frequency Variability

For the binomial model,

$$
E[N]=np,
\qquad
\operatorname{Var}(N)=np(1-p).
$$

For the Poisson model,

$$
E[N]=\lambda,
\qquad
\operatorname{Var}(N)=\lambda.
$$

For the negative binomial model,

$$
E[N]=\frac{kq}{p},
\qquad
\operatorname{Var}(N)=\frac{kq}{p^2}.
$$

Thus, the binomial model has $\operatorname{Var}(N)<E[N],$

while the Poisson and negative binomial models have variance equal to or greater than their mean.

## Interpretation of Frequency Variability

The three models represent different patterns of claim-count variability.

**Binomial:** fixed number of potential claim events and limited frequency variability.

**Poisson:** no fixed upper limit, with frequency variance equal to the mean.

**Negative binomial:** no fixed upper limit and greater frequency variability than the Poisson model.

The choice of frequency model should reflect the underlying claim-generation mechanism.

## A Fixed Upper Limit on Claims

A defining feature of the compound binomial model is

$$
N\leq n.
$$

Consequently, the portfolio can contain at most $n$ claims.

This is natural when there are $n$ policies and each policy can produce at most one claim.

In contrast, Poisson and negative binomial models have no finite upper limit on $N$.

## The Role of the Claim Probability

For

$$
N\sim\operatorname{Binomial}(n,p),
$$

the expected number of claims is

$$
E[N]=np.
$$

Increasing $p$ increases the expected frequency.

However,$\operatorname{Var}(N)=np(1-p),$ so frequency variability is also affected by $p$.

The variance is largest when $p=0.5$.

## Aggregate Loss Depends on Frequency and Severity

For the compound binomial model,

$$
E[S]=npE[X].
$$

The expected aggregate loss therefore depends on:

* the number of potential claims $n$;
* the probability of a claim $p$;
* the expected claim amount $E[X]$.

Similarly,

$$
\operatorname{Var}(S)
=
E[N]\operatorname{Var}(X)
+
\operatorname{Var}(N)\{E[X]\}^2.
$$

Thus, aggregate risk comes from both **claim severity** and **claim frequency**.

## Binomial to Poisson as an Approximation

The binomial frequency model approaches a Poisson model when

$$
n\to\infty,
\qquad
p\to0,
\qquad
np\to\lambda.
$$

Then

$$
\operatorname{Binomial}(n,p)
\approx
\operatorname{Poisson}(\lambda).
$$

This provides a useful connection between the compound binomial and compound Poisson models.

The Poisson model can therefore be viewed as an approximation when there are many potential claim events with small individual claim probabilities.


## Compound Binomial: Key Results

For

$$
N\sim\operatorname{Binomial}(n,p),
$$

and

$$
S=\sum_{i=1}^{N}X_i,
$$

with $N$ independent of the i.i.d. claim amounts $X_i$,

$$
E[S]=np\,m_1,\qquad \operatorname{Var}(S)=npm_2-np^2m_1^2.
$$


The compound binomial model is particularly appropriate when there is a **fixed number of potential claim events**, with each event producing at most one claim.

## Three Compound Models

The main distinction among the three special compound models is their **frequency distribution**.

$$
\begin{array}{c|c|c}
\text{Model} & E[N] & \operatorname{Var}(N)\\
\hline
\text{Binomial} & np & np(1-p)\\
\text{Poisson} & \lambda & \lambda\\
\text{Negative Binomial} & \dfrac{kq}{p} & \dfrac{kq}{p^2}
\end{array}
$$

The frequency assumption determines the variability of the aggregate loss.



# Aggregate Claims with Reinsurance

## Next: Aggregate Claims with Reinsurance

We now extend the collective risk model to incorporate **reinsurance**.

The next section considers aggregate claim distributions under:

* **Proportional reinsurance**
* **Individual excess of loss reinsurance**

We will examine how reinsurance changes the **individual claim amounts** and, consequently, the **aggregate claim distribution**.







## Proportional Reinsurance

Let $S$ denote the total aggregate claims from a risk during a given period.

Let $S_I$ and $S_R$ denote the aggregate claims paid by the **insurer** and **reinsurer**, respectively.

Then

$$
\boxed{S=S_I+S_R}.
$$

Under proportional reinsurance, the insurer retains a fixed proportion $\alpha$ of every claim.

## Proportional Reinsurance

For each claim $X_i$,

$$
Y_i=\alpha X_i,
\qquad
Z_i=(1-\alpha)X_i.
$$

Therefore,

$$
S_I=\sum_{i=1}^N\alpha X_i=\alpha S
$$

and

$$
S_R=\sum_{i=1}^N(1-\alpha)X_i=(1-\alpha)S.
$$

## Key Features

* Both the insurer and reinsurer contribute to **every claim**.
* The insurer retains the same proportion $\alpha$ of every claim.
* The reinsurer pays the remaining proportion $1-\alpha$.
* Both parties have **unlimited liability** unless a separate cap is imposed.

Thus,

$$
\boxed{S_I=\alpha S,\qquad S_R=(1-\alpha)S}.
$$

## Worked Example

Suppose aggregate claims follow a compound Poisson distribution with

$$
N\sim\operatorname{Poisson}(10)
$$

and individual claim amounts

$$
X\sim Pa(4,1).
$$

The insurer purchases proportional reinsurance with retention $\alpha=0.8.$

Find:

1. the distributions of $S_I$ and $S_R$;
2. their means and variances;
3. compare $\operatorname{Var}(S_I)+\operatorname{Var}(S_R)$ with $\operatorname{Var}(S)$.


## Solution: Set Up the Reinsurance Model

The original aggregate claim is

$$
S=\sum_{i=1}^{N}X_i.
$$

Under proportional reinsurance with $\alpha=0.8$,

$$
S_I=0.8S,
\qquad
S_R=0.2S.
$$

Equivalently,

$$
S_I=\sum_{i=1}^{N}0.8X_i,
\qquad
S_R=\sum_{i=1}^{N}0.2X_i.
$$

## Solution: Claim Amounts After Reinsurance

Define the insurer's payment per claim as

$$
Y_i=0.8X_i.
$$

Define the reinsurer's payment per claim as

$$
Z_i=0.2X_i.
$$

Then

$$
S_I=\sum_{i=1}^{N}Y_i,
\qquad
S_R=\sum_{i=1}^{N}Z_i.
$$

Both are therefore compound Poisson distributions.

## Solution: Distribution of the Insurer's Claim

Recall the scaling property:

$$
X\sim Pa(\alpha,\lambda)
\quad\Longrightarrow\quad
kX\sim Pa(\alpha,k\lambda).
$$

Since

$$
X_i\sim Pa(4,1)
$$

and

$$
Y_i=0.8X_i,
$$

we obtain

$$
\boxed{Y_i\sim Pa(4,0.8)}.
$$

## Solution: Distribution of the Reinsurer's Claim

Similarly,

$$
Z_i=0.2X_i.
$$

Therefore,

$$
\boxed{Z_i\sim Pa(4,0.2)}.
$$

Hence,

$$
\boxed{
S_I\sim\mathcal{CP}(10,Pa(4,0.8))
}
$$

and

$$
\boxed{
S_R\sim\mathcal{CP}(10,Pa(4,0.2))
}.
$$

## Solution: Moments of the Original Severity

For

$$
X\sim Pa(4,1),
$$

the mean is

$$
E[X]=\frac{\lambda}{\alpha-1}
=\frac{1}{4-1}
=\frac{1}{3}.
$$

The second moment is

$$
E[X^2]
=
\frac{\Gamma(\alpha-2)\Gamma(3)}
{\Gamma(\alpha)}
\lambda^2.
$$

Thus, $E[X^2]=\frac{\Gamma(2)\Gamma(3)}{\Gamma(4)}=\frac{1}{3}.$

## Solution: Original Aggregate Moments

For a compound Poisson model,

$$
E[S]=\lambda E[X]
$$

and

$$
\operatorname{Var}(S)=\lambda E[X^2].
$$

Here $\lambda=10$, so

$$
E[S]
=
10\left(\frac{1}{3}\right)
=
\boxed{\frac{10}{3}}.
$$

## Solution: Original Aggregate Variance

Using

$$
\operatorname{Var}(S)=\lambda E[X^2],
$$

we obtain

$$
\operatorname{Var}(S)
=
10\left(\frac{1}{3}\right).
$$

Therefore,

$$
\boxed{\operatorname{Var}(S)=\frac{10}{3}}.
$$

## Solution: Insurer's Mean

Since

$$
S_I=0.8S,
$$

we can use the scaling property of expectation:

$$
E[S_I]
=
0.8E[S].
$$

Therefore,

$$
E[S_I]
=
0.8\left(\frac{10}{3}\right)
=
\boxed{\frac{8}{3}}.
$$

## Solution: Insurer's Variance

For a constant $c$,

$$
\operatorname{Var}(cS)=c^2\operatorname{Var}(S).
$$

Therefore,

$$
\operatorname{Var}(S_I)
=
(0.8)^2\operatorname{Var}(S).
$$

Hence,

$$
\operatorname{Var}(S_I)
=
(0.8)^2\left(\frac{10}{3}\right)
=
\boxed{\frac{32}{15}}.
$$

## Solution: Reinsurer's Mean

Similarly,

$$
S_R=0.2S.
$$

Therefore,

$$
E[S_R]
=
0.2E[S]
=
0.2\left(\frac{10}{3}\right).
$$

Thus,

$$
\boxed{E[S_R]=\frac{2}{3}}.
$$

## Solution: Reinsurer's Variance

Using the variance scaling property,

$$
\operatorname{Var}(S_R)
=
(0.2)^2\operatorname{Var}(S).
$$

Therefore,

$$
\operatorname{Var}(S_R)
=
(0.2)^2\left(\frac{10}{3}\right)
=
\boxed{\frac{2}{15}}.
$$

## Solution: Compare the Variances

The sum of the two individual variances is

$$
\operatorname{Var}(S_I)+\operatorname{Var}(S_R)
=
\frac{32}{15}+\frac{2}{15}
=
\boxed{\frac{34}{15}}.
$$

The original aggregate variance is

$$
\operatorname{Var}(S)=\frac{10}{3}
=\frac{50}{15}.
$$

Therefore,

$$
\boxed{
\frac{34}{15}<\frac{50}{15}
}.
$$

## Solution: Why Are They Different?

We know that

$$
S=S_I+S_R.
$$

Thus,

$$
\operatorname{Var}(S)
=
\operatorname{Var}(S_I)
+
\operatorname{Var}(S_R)
+
2\operatorname{Cov}(S_I,S_R).
$$

The covariance term is important because

$$
S_I=0.8S,
\qquad
S_R=0.2S.
$$

## Solution: The Covariance

We have

$$
\operatorname{Cov}(S_I,S_R)
=
\operatorname{Cov}(0.8S,0.2S).
$$

Therefore,

$$
\operatorname{Cov}(S_I,S_R)
=
0.8(0.2)\operatorname{Var}(S).
$$

Hence,

$$
\operatorname{Cov}(S_I,S_R)
=
0.16\left(\frac{10}{3}\right)
=
\boxed{\frac{8}{15}}.
$$

So the variance decomposition becomes $\frac{32}{15}+\frac{2}{15}+2\left(\frac{8}{15}\right)=\frac{50}{15}=\frac{10}{3}.$


## Key Takeaways

Under proportional reinsurance,

$$
\boxed{S_I=\alpha S,\qquad S_R=(1-\alpha)S}.
$$

Therefore,

$$
E[S_I]=\alpha E[S],
\qquad
E[S_R]=(1-\alpha)E[S].
$$

and

$$
\operatorname{Var}(S_I)=\alpha^2\operatorname{Var}(S),
\qquad
\operatorname{Var}(S_R)=(1-\alpha)^2\operatorname{Var}(S).
$$

The insurer and reinsurer payments are **dependent**, so

$$
\operatorname{Var}(S_I)+\operatorname{Var}(S_R)
\neq
\operatorname{Var}(S).
$$










