---
title: "Solutions to Mock Examination: 2020"
format:
  pdf:
    toc: false
number-sections: false
---



1. Claims occur on a general insurance portfolio independently and at random. Each claim is classified as being of Type A or Type B. Type A claim amounts are distributed $\operatorname{Pa}(4,500)$ and Type B claim amounts are distributed $\operatorname{Pa}(6,2000)$. It is known that 80\% of all claims are of Type A.

   Let $X$ denote a claim chosen at random from the portfolio.

   1. Calculate $\Pr(X > 1500)$.

   2. Calculate $E[X]$ and $\operatorname{Var}(X)$.

   3. Suppose $Y$ has a Pareto distribution with the same mean and variance as $X$. Calculate $\Pr(Y > 1500)$.

   4. Comment on the difference between the answers in parts 1 and 3.
   


```{r}
# Parameters
p_A <- 0.80
p_B <- 0.20

a_A <- 4
lambda_A <- 500

a_B <- 6
lambda_B <- 2000

d <- 1500

# Pareto Type II survival function
pareto_tail <- function(x, a, lambda) {
  (lambda / (lambda + x))^a
}

# 1. P(X > 1500)
tail_A <- pareto_tail(d, a_A, lambda_A)
tail_B <- pareto_tail(d, a_B, lambda_B)

p_tail_X <- p_A * tail_A + p_B * tail_B

# 2. E[X] and Var(X)
pareto_mean <- function(a, lambda) {
  lambda / (a - 1)
}

pareto_var <- function(a, lambda) {
  a * lambda^2 / ((a - 1)^2 * (a - 2))
}

mean_A <- pareto_mean(a_A, lambda_A)
mean_B <- pareto_mean(a_B, lambda_B)

var_A <- pareto_var(a_A, lambda_A)
var_B <- pareto_var(a_B, lambda_B)

mean_X <- p_A * mean_A + p_B * mean_B

# E[X^2] = Var(X) + E[X]^2
EX2_A <- var_A + mean_A^2
EX2_B <- var_B + mean_B^2

EX2_X <- p_A * EX2_A + p_B * EX2_B

var_X <- EX2_X - mean_X^2

# 3. Pareto distribution Y with the same mean and variance
# For Pareto Type II:
# E[Y] = lambda_Y / (a_Y - 1)
# Var(Y) = a_Y lambda_Y^2 / ((a_Y - 1)^2 (a_Y - 2))

a_Y <- 2*(var_X/mean_X^2)*(1/((var_X/mean_X^2)-1))
lambda_Y <- mean_X * (a_Y - 1)

p_tail_Y <- pareto_tail(d, a_Y, lambda_Y)

# Difference
difference <- p_tail_Y - p_tail_X

# Display results
p_tail_X
mean_X
var_X
a_Y
lambda_Y
p_tail_Y
difference
```

1. For Type A claims,

   $$
   \Pr(X_A>1500)
   =
   \left(\frac{500}{500+1500}\right)^4
   =
   `r round(tail_A, 6)`.
   $$

   For Type B claims,

   $$
   \Pr(X_B>1500)
   =
   \left(\frac{2000}{2000+1500}\right)^6
   =
   `r round(tail_B, 6)`.
   $$

   Since 80% of claims are Type A,

   $$
   \begin{aligned}
   \Pr(X>1500)
   &=0.8\Pr(X_A>1500)+0.2\Pr(X_B>1500)\\
   &=0.8(`r round(tail_A, 6)`)+0.2(`r round(tail_B, 6)`)\\
   &=`r round(p_tail_X, 6)`.
   \end{aligned}
   $$

2. For a Pareto Type II distribution,

   $$
   E[X]=\frac{\lambda}{a-1},
   \qquad
   \operatorname{Var}(X)
   =
   \frac{a\lambda^2}{(a-1)^2(a-2)}.
   $$

   Therefore, for Type A,

   $$
   E[X_A]
   =
   \frac{500}{4-1}
   =
   `r round(mean_A, 4)`,
   $$

   and

   $$
   \operatorname{Var}(X_A)
   =
   \frac{4(500)^2}{(4-1)^2(4-2)}
   =
   `r round(var_A, 4)`.
   $$

   For Type B,

   $$
   E[X_B]
   =
   \frac{2000}{6-1}
   =
   `r round(mean_B, 4)`,
   $$

   and

   $$
   \operatorname{Var}(X_B)
   =
   \frac{6(2000)^2}{(6-1)^2(6-2)}
   =
   `r round(var_B, 4)`.
   $$

   Hence,

   $$
   \begin{aligned}
   E[X]
   &=0.8E[X_A]+0.2E[X_B]\\
   &=0.8(`r round(mean_A, 4)`)+0.2(`r round(mean_B, 4)`)\\
   &=`r round(mean_X, 4)`.
   \end{aligned}
   $$

   To calculate the variance of the mixture, first calculate the second moment:

   $$
   \begin{aligned}
   E[X^2]
   &=0.8E[X_A^2]+0.2E[X_B^2]\\
   &=0.8\left(\operatorname{Var}(X_A)+E[X_A]^2\right)
   +0.2\left(\operatorname{Var}(X_B)+E[X_B]^2\right).
   \end{aligned}
   $$

   Thus,

   $$
   E[X^2]=`r round(EX2_X, 4)`,
   $$

   and

   $$
   \begin{aligned}
   \operatorname{Var}(X)
   &=E[X^2]-\{E[X]\}^2\\
   &=`r round(EX2_X, 4)`-(`r round(mean_X, 4)`)^2\\
   &=`r round(var_X, 4)`.
   \end{aligned}
   $$

3. Let $Y$ have a Pareto distribution with the same mean and variance as $X$. Write its parameters as $a_Y$ and $\lambda_Y$.

   For a Pareto Type II distribution,

   $$
   E[Y]=\frac{\lambda_Y}{a_Y-1},
   $$

   and

   $$
   \operatorname{Var}(Y)
   =
   \frac{a_Y\lambda_Y^2}
   {(a_Y-1)^2(a_Y-2)}.
   $$

   Matching the mean and variance gives

   $$
   a_Y
   =
   2+\frac{E[X]^2}{\operatorname{Var}(X)}
   =
   2+\frac{(`r round(mean_X, 4)`)^2}{`r round(var_X, 4)`}
   =
   `r round(a_Y, 6)`.
   $$

   Then

   $$
   \lambda_Y
   =
   E[X](a_Y-1)
   =
   (`r round(mean_X, 4)`)(
   `r round(a_Y, 6)`-1)
   =
   `r round(lambda_Y, 6)`.
   $$

   Therefore,

   $$
   \begin{aligned}
   \Pr(Y>1500)
   &=
   \left(
   \frac{\lambda_Y}{\lambda_Y+1500}
   \right)^{a_Y}\\
   &=
   \left(
   \frac{`r round(lambda_Y, 6)`}
   {`r round(lambda_Y, 6)`+1500}
   \right)^{`r round(a_Y, 6)`}\\
   &=`r round(p_tail_Y, 6)`.
   \end{aligned}
   $$

4. The two probabilities are

   $$
   \Pr(X>1500)=`r round(p_tail_X, 6)`
   $$

   and

   $$
   \Pr(Y>1500)=`r round(p_tail_Y, 6)`.
   $$

   Thus,

   $$
   \Pr(Y>1500)-\Pr(X>1500)
   =
   `r round(difference, 6)`.
   $$

   Although $X$ and $Y$ have the **same mean and variance**, they do not have the same distribution. The distribution of $X$ is a **mixture of two Pareto distributions**, whereas $Y$ is a **single Pareto distribution**. Consequently, matching only the first two moments does not guarantee that the tail probabilities will be the same.
   
   
   
<!-- A comment or text block -->   
   
---
   
2. Consider a portfolio of motor insurance policies. In the event of an accident, the cost of repairs to a car has a Pareto distribution with parameters $\alpha$ and $\lambda$. A deductible of 50 is applied to all claims, and a claim is made whenever the repair cost exceeds the deductible. A sample of 200 claims has a mean of 100 and a standard deviation of 150.**

    1. Using the method of moments, estimate $\alpha$ and $\lambda$.
    2. Estimate the proportion of accidents that do not result in a claim.
    3. The insurance company arranges excess-of-loss reinsurance with another insurance company to reduce the mean amount it pays per claim to 70. Calculate the retention limit needed to achieve this.

**R code**

```{r}
# Given sample information
d <- 50
mean_claim <- 100
sd_claim <- 150
var_claim <- sd_claim^2

# ------------------------------------------------------------
# 1. Method of moments
#
# Given X | X > d, the excess claim Y = X - d has
#
# E[Y] = (lambda + d)/(alpha - 1)
#
# Var(Y) = alpha (lambda + d)^2 /
#          [(alpha - 1)^2 (alpha - 2)]
#
# Therefore:
# Var(Y) / E[Y]^2 = alpha / (alpha - 2)
# ------------------------------------------------------------

alpha_hat <- 2 * var_claim / (var_claim - mean_claim^2)

lambda_plus_d <- mean_claim * (alpha_hat - 1)

lambda_hat <- lambda_plus_d - d

# ------------------------------------------------------------
# 2. Probability that an accident does not result in a claim
#
# P(X <= d) = 1 - [lambda/(lambda+d)]^alpha
# ------------------------------------------------------------

p_no_claim <- 1 - (lambda_hat / (lambda_hat + d))^alpha_hat

# ------------------------------------------------------------
# 3. XOL retention limit
#
# Let Y = X - d | X > d.
#
# E[min(Y, M)]
#   = integral_0^M P(Y > y) dy
#   = (lambda+d)/(alpha-1) *
#     [1 - ((lambda+d)/(lambda+d+M))^(alpha-1)]
#
# Set this equal to 70 and solve for M.
# ------------------------------------------------------------

target_mean <- 70

M <- (lambda_plus_d) *
  ((1 - target_mean / mean_claim)^(-1 / (alpha_hat - 1)) - 1)

# ------------------------------------------------------------
# Display results
# ------------------------------------------------------------

alpha_hat
lambda_hat
p_no_claim
M

# Optional: verify the retained mean
retained_mean <- mean_claim *
  (1 - (lambda_plus_d / (lambda_plus_d + M))^(alpha_hat - 1))

retained_mean
```

1. Let $X$ denote the original repair cost, with the two-parameter Pareto distribution

   $$
   P(X>x)=\left(\frac{\lambda}{\lambda+x}\right)^\alpha.
   $$

   Because the sample consists only of claims, the observations are from the **excess distribution**

   $$
   Y=X-50\mid X>50.
   $$

   By the threshold stability of the Pareto distribution,

   $$
   P(Y>y)
   =
   \left(\frac{\lambda+50}{\lambda+50+y}\right)^\alpha.
   $$

   Thus, $Y$ is again Pareto in the same parameterization, with scale parameter $\lambda+50$.

   The sample moments are

   $$
   E[Y]=100,
   \qquad
   \operatorname{Var}(Y)=150^2=22500.
   $$

   For the excess distribution,

   $$
   E[Y]=\frac{\lambda+50}{\alpha-1},
   $$

   and

   $$
   \operatorname{Var}(Y)
   =
   \frac{\alpha(\lambda+50)^2}
   {(\alpha-1)^2(\alpha-2)}.
   $$

   Dividing the variance equation by the square of the mean gives

   $$
   \frac{\operatorname{Var}(Y)}{E[Y]^2}
   =
   \frac{\alpha}{\alpha-2}.
   $$

   Hence,

   $$
   \frac{22500}{100^2}
   =
   \frac{\alpha}{\alpha-2}.
   $$

   Solving for $\alpha$ gives

   $$
   \widehat{\alpha}
   =
   `r alpha_hat`.
   $$

   Using the first moment,

   $$
   100
   =
   \frac{\lambda+50}{`r alpha_hat`-1}.
   $$

   Therefore,

   $$
   \widehat{\lambda}+50
   =
   100(`r alpha_hat`-1)
   =
   `r lambda_plus_d`,
   $$

   and hence

   $$
   \boxed{\widehat{\alpha}=`r alpha_hat`,
   \qquad
   \widehat{\lambda}=`r lambda_hat`.}
   $$

2. An accident does not result in a claim when the repair cost does not exceed the deductible. Therefore,

   $$
   P(\text{no claim})
   =
   P(X\leq50)
   =
   F_X(50).
   $$

   Using the fitted Pareto distribution,

   $$
   P(X\leq50)
   =
   1-
   \left(
   \frac{\widehat{\lambda}}
   {\widehat{\lambda}+50}
   \right)^{\widehat{\alpha}}.
   $$

   Substituting the estimates,

   $$
   P(\text{no claim})
   =
   `r p_no_claim`.
   $$

   Thus, the estimated proportion of accidents that do not result in a claim is

   $$
   \boxed{`r 100*p_no_claim`\%.}
   $$

<!-- A comment or text block -->

3. For an excess-of-loss retention limit $M$, the insurance company pays

   $$
   \min(Y,M),
   $$

   where $Y=X-50\mid X>50$ is the amount of the claim above the deductible.

   The survival function of $Y$ is

   $$
   P(Y>y)
   =
   \left(
   \frac{\lambda+50}
   {\lambda+50+y}
   \right)^\alpha.
   $$

   The limited expected value is therefore

   $$
   E[\min(Y,M)]
   =
   \int_0^M P(Y>y)\,dy.
   $$

   Evaluating the integral,

   $$
   E[\min(Y,M)]
   =
   \frac{\lambda+50}{\alpha-1}
   \left[
   1-
   \left(
   \frac{\lambda+50}
   {\lambda+50+M}
   \right)^{\alpha-1}
   \right].
   $$

  The original mean claim amount is 100, and the required mean amount paid by the insurer is 70. Hence,

   $$
   70
   =
   100
   \left[
   1-
   \left(
   \frac{`r lambda_plus_d`}
   {`r lambda_plus_d`+M}
   \right)^{`r alpha_hat`-1}
   \right].
   $$

   Rearranging,

   $$
   \left(
   \frac{`r lambda_plus_d`}
   {`r lambda_plus_d`+M}
   \right)^{`r alpha_hat`-1}
   =
   1-\frac{70}{100}.
   $$

   Therefore,

   $$
   M
   =
   `r lambda_plus_d`
   \left[
   \left(1-\frac{70}{100}\right)^{-1/(`r alpha_hat`-1)}
   -1
   \right].
   $$

   Thus, the required retention limit is

   $$
   \boxed{M=`r M`.}
   $$

  As a check, substituting this retention limit into the limited expected value formula gives

   $$
   E[\min(Y,M)]
   =
   `r retained_mean`,
   $$

   which is the required mean amount of 70, up to numerical precision.
   


---

3. Consider aggregate claims $S$ with a compound Poisson distribution $\mathcal{CP}(10,F_X)$, where the individual claim amount $X$ takes values $2$, $4$, $8$, and $10$ with probabilities $0.4$, $0.3$, $0.2$, and $0.1$, respectively. The insurer enters into an excess-of-loss reinsurance contract with retention level $M=6$.

    1. Calculate $E[X]$ and $E[X^2]$.
    2. Calculate $E[S]$ and $\operatorname{Var}(S)$.
    3. Determine the distributions of the aggregate amounts paid by the insurer, $S_I$, and the reinsurer, $S_R$, and calculate $E[S_I]$ and $E[S_R]$.



```{r}
# Parameters
lambda <- 10
M <- 6

x <- c(2, 4, 8, 10)
p <- c(0.4, 0.3, 0.2, 0.1)

# 1. Moments of X
EX <- sum(x * p)
EX2 <- sum(x^2 * p)

# 2. Moments of aggregate claims S
ES <- lambda * EX
VarS <- lambda * EX2

# 3. XOL payments
XI <- pmin(x, M)          # insurer payment
XR <- pmax(x - M, 0)      # reinsurer payment

# Aggregate means
EXI <- sum(XI * p)
EXR <- sum(XR * p)

ESI <- lambda * EXI
ESR <- lambda * EXR

# Display results
EX
EX2
ES
VarS
data.frame(x, p, XI, XR)
EXI
EXR
ESI
ESR
```

1. For the individual claim amount $X$, we have

   $$
   P(X=2)=0.4,\qquad
   P(X=4)=0.3,\qquad
   P(X=8)=0.2,\qquad
   P(X=10)=0.1.
   $$

   Therefore,

   $$
   E[X]
   =\sum_x xP(X=x)
   =2(0.4)+4(0.3)+8(0.2)+10(0.1).
   $$

   Using the R code,

   `r EX <- sum(x*p); EX`

   Hence,

   $$
   \boxed{E[X]=`r EX`}.
   $$

   The second moment is

   $$
   E[X^2]
   =\sum_x x^2P(X=x)
   =2^2(0.4)+4^2(0.3)+8^2(0.2)+10^2(0.1).
   $$

   Using R,

   `r EX2 <- sum(x^2*p); EX2`

   Thus,

   $$
   \boxed{E[X^2]=`r EX2`}.
   $$

2. Since $S$ is compound Poisson with Poisson parameter $\lambda=10$,

   $$
   E[S]=\lambda E[X].
   $$

   Therefore,

   $$
   E[S]
   =`r lambda`(`r EX`)
   =`r ES`.
   $$

   Hence,

   $$
   \boxed{E[S]=`r ES`}.
   $$

   For a compound Poisson distribution,

   $$
   \operatorname{Var}(S)=\lambda E[X^2].
   $$

   Therefore,

   $$
   \operatorname{Var}(S)
   =`r lambda`(`r EX2`)
   =`r VarS`.
   $$

   Hence,

   $$
   \boxed{\operatorname{Var}(S)=`r VarS`}.
   $$

3. Under the excess-of-loss contract with retention $M=6$, the insurer pays

   $$
   X_I=\min(X,6),
   $$

   while the reinsurer pays

   $$
   X_R=(X-6)_+=\max(X-6,0).
   $$

   The claim-by-claim payments are therefore

   $$
   \begin{array}{c|cccc}
   X & 2 & 4 & 8 & 10\\
   \hline
   X_I & 2 & 4 & 6 & 6\\
   X_R & 0 & 0 & 2 & 4
   \end{array}
   $$

   Thus, the insurer's payment per original claim has distribution

   $$
   P(X_I=2)=0.4,
   $$

   $$
   P(X_I=4)=0.3,
   $$

   and

   $$
   P(X_I=6)=0.2+0.1=0.3.
   $$

   Hence,

   $$
   X_I=
   \begin{cases}
   2,&P=0.4,\\
   4,&P=0.3,\\
   6,&P=0.3.
   \end{cases}
   $$

   The insurer's aggregate payment is

   $$
   S_I=\sum_{i=1}^{N}\min(X_i,6),
   \qquad N\sim\operatorname{Poisson}(10).
   $$

   Therefore, $S_I$ has a **compound Poisson distribution** with frequency parameter $10$ and severity distribution given in the previous step:

   $$
   \boxed{
   S_I\sim\mathcal{CP}\left(
   10,
   \begin{cases}
   2&0.4\\
   4&0.3\\
   6&0.3
   \end{cases}
   \right)
   }.
   $$

   The expected insurer payment per claim is

   $$
   E[X_I]
   =2(0.4)+4(0.3)+6(0.3).
   $$

   Using R,

   `r EXI <- sum(pmin(x,M)*p); EXI`

   Thus,

   $$
   E[X_I]=`r EXI`.
   $$

   Consequently,

   $$
   E[S_I]
   =10E[X_I]
   =`r ESI`.
   $$

   Hence,

   $$
   \boxed{E[S_I]=`r ESI`}.
   $$

   For the reinsurer, the payment per claim is

   $$
   X_R=(X-6)_+.
   $$

   Therefore,

   $$
   P(X_R=0)=0.4+0.3=0.7,
   $$

   $$
   P(X_R=2)=0.2,
   $$

   and

   $$
   P(X_R=4)=0.1.
   $$

   Hence,

   $$
   X_R=
   \begin{cases}
   0,&P=0.7,\\
   2,&P=0.2,\\
   4,&P=0.1.
   \end{cases}
   $$

   The reinsurer's aggregate payment can therefore be written as

$$
S_R=\sum_{i=1}^{N}(X_i-6)_+,
\qquad N\sim\operatorname{Poisson}(10).
$$

   Thus, strictly including the zero payments, $S_R$ is a compound Poisson distribution with severity distribution

$$
\boxed{
X_R=
\begin{cases}
0&0.7\\
2&0.2\\
4&0.1.
\end{cases}
}
$$

   Equivalently, the positive reinsurer payments occur by **Poisson thinning**, so the number of claims producing a positive reinsurance payment is

$$
N_R\sim\operatorname{Poisson}\left(10(0.2+0.1)\right)
=\operatorname{Poisson}(3),
$$

   with positive payment distribution

$$
P(X_R=2\mid X_R>0)=\frac{0.2}{0.3}=\frac{2}{3},
$$

   and

$$
P(X_R=4\mid X_R>0)=\frac{0.1}{0.3}=\frac{1}{3}.
$$

   The expected reinsurer payment per original claim is

$$
E[X_R]
=0(0.7)+2(0.2)+4(0.1).
$$

   Using R, `r EXR <- sum(pmax(x-M,0)*p); EXR`

   Hence,

$$
E[X_R]=`r EXR`.
$$

   Therefore,

$$
E[S_R]
=10E[X_R]
=`r ESR`.
$$

   Thus,

$$
\boxed{E[S_R]=`r ESR`}.
$$

   As a useful check, the insurer and reinsurer payments must add to the original claim amount:

$$
X=X_I+X_R.
$$

   Consequently,

$$
S=S_I+S_R,
$$

   and therefore their means satisfy

$$
E[S]=E[S_I]+E[S_R].
$$

   Indeed,

$$
`r ESI`+`r ESR`=`r ESI + ESR`,
$$

   which agrees with

$$
E[S]=`r ES`.
$$

   This confirms that the original aggregate expected claim amount is divided between the insurer and reinsurer as required by the excess-of-loss contract.











---


4. The number of claims $N$ during a fixed time period has the distribution
$$P(N=0)=0.4,\qquad P(N=1)=0.3,\qquad P(N=2)=0.2,\qquad P(N=3)=0.1.$$
Claim amounts are uniformly distributed on $(0,200)$, and the number of claims is independent of the claim amounts.

   1. Find the mean and variance of the aggregate claims during the fixed time period.
   2. Suppose a policy deductible of $10$ applies to each claim. Find the expected insurer payout.

```{r}
# Parameters
prob_N <- c(0.4, 0.3, 0.2, 0.1)
N_values <- 0:3

# Claim amount X ~ Uniform(0, 200)
a <- 0
b <- 200

# Moments of N
EN <- sum(N_values * prob_N)
EN2 <- sum(N_values^2 * prob_N)
VarN <- EN2 - EN^2

# Moments of X
EX <- (a + b) / 2
VarX <- (b - a)^2 / 12
EX2 <- VarX + EX^2

# Aggregate claims S = sum_{i=1}^N X_i
ES <- EN * EX
VarS <- EN * VarX + VarN * EX^2

# Deductible d = 10
d <- 10

# Insurer payment per claim Y = (X - d)_+
# E[Y] = E[X] - d + E[(d-X)_+]
# For X ~ Uniform(0,200):
EY <- (b - d)^2 / (2 * (b - a))

# Expected aggregate insurer payout
ES_deductible <- EN * EY

# Display results
EN
VarN
EX
VarX
ES
VarS
EY
ES_deductible
```

1. Let the aggregate claims be
   $S=\sum_{i=1}^{N}X_i,$
   where $X_i$ denotes the amount of the $i$th claim. First calculate the moments of the claim count:

   $$E[N]=0(0.4)+1(0.3)+2(0.2)+3(0.1)
   =`r EN`.$$

   Also,
   $$E[N^2]=0^2(0.4)+1^2(0.3)+2^2(0.2)+3^2(0.1)
   =`r EN2`.$$

   Hence,
   $$\operatorname{Var}(N)=E[N^2]-[E(N)]^2
   =`r EN2`-(`r EN`)^2
   =`r VarN`.$$


   Since $X\sim\operatorname{Uniform}(0,200)$,
   $E[X]=\frac{0+200}{2}=`r EX`,$
   and
   $\operatorname{Var}(X)=\frac{(200-0)^2}{12}=`r VarX`.$

   Using the standard random-sum formulas, and the independence of $N$ and the claim amounts,
   $E[S]=E[N]E[X].$
   Therefore,
   $E[S]=`r EN`\times `r EX`=`r ES`.$

   For the variance of a random sum,

   $$\operatorname{Var}(S)
   =E[N]\operatorname{Var}(X)
   +\operatorname{Var}(N)[E(X)]^2.$$
   Thus,
   $$\begin{aligned}
   \operatorname{Var}(S)
   &=(`r EN`)(`r VarX`)
   +(`r VarN`)(`r EX`^2)\\
   &=`r VarS`.
   \end{aligned}$$


2. Now suppose a deductible of $10$ applies to each claim. The insurer's payment for an individual claim is
   $Y=(X-10)_+=\max(X-10,0).$
   Since $X\sim\operatorname{Uniform}(0,200)$,

   $$E[Y]
   =\frac{1}{200}\int_{10}^{200}(x-10)\,dx
   =\frac{(200-10)^2}{2(200)}
   =`r EY`.$$

   Let the aggregate insurer payout after the deductible be
   $S_D=\sum_{i=1}^{N}(X_i-10)_+.$
   By independence of $N$ and the claim amounts,
   $E[S_D]=E[N]E[(X-10)_+].$
   Therefore,
   $E[S_D]=(`r EN`)(`r EY`)=`r ES_deductible`.$

   Hence, the answers are
   $\boxed{E[S]=`r ES`}$
   and
   $\boxed{\operatorname{Var}(S)=`r VarS`}.$
   With a deductible of $10$, the expected insurer payout is
   $\boxed{E[S_D]=`r ES_deductible`}.$

   